---
title: "Analysis"
author: "Daniel Fuller"
output:
      html_document:
        keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(rstatix)
library(haven)
library(kableExtra)
library(lme4)
```

## Analysis for Sleep, PA Cancer

```{r}
data <- read_csv("/Users/dlf545/Documents/Ryan Collins Thesis/sleep_pa_cancer/analysis_data.csv")
```

## Quick example regression based on research question 1

To examine if higher levels of MVPA at baseline are associated with better subjective and objective cognitive function outcomes at T4 (1-year post-diagnosis) in the first year of cancer diagnosis in women with breast cancer.

* Outcome is subjective cognitive functionning (FACTCog_TOTAL_Score)
* Predictors are minutes of MVPA, Sleep, T1_Caffeine, and BMI

## Plot to check variation over time

```{r}
plot1 <- ggplot(data, aes(time, FACTCog_TOTAL_Score, group=p_id, color=p_id)) +
  geom_smooth(method="lm", se=F) +
  geom_jitter(size=1) +
  theme(legend.position="none") 
plot(plot1)
```

This plot shows the individual trajectories of participants over time for the subjective cognitive functionning (FACTCog_TOTAL_Score) measure. Clearly differences in individuals and slops (we will deal with slopes late).

### Example regression

```{r}
model1 <- lmer(FACTCog_TOTAL_Score ~ sleep_minutes + mvpa_minutes + T1_Caffeine + BMI + (1 | p_id), data = data)
summary(model1)
```




