---
title: "Data Wrangling"
author: "Daniel Fuller"
date: "26/11/2021"
output:
      html_document:
        keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(rstatix)
library(haven)
```

# Data Wrangling for Sleep, PA, and Cancer study

```{r}
sleep_data_temp <- NULL

sleep_data_temp <- dir(path = "Participant Data/", full.names = TRUE, pattern = "*.csv", recursive = TRUE) %>%
   purrr::map(function(i){
     dfhx <- read_csv(i, col_types = cols(.default = col_character()))
     dfhx$p_id <- stringr::str_sub(i, 19, 21)
     dfhx$time <- stringr::str_sub(i, 23, 24)
     sleep_data_temp <- bind_rows(sleep_data_temp, dfhx)
   })

merged_data <- bind_rows(sleep_data_temp)

glimpse(merged_data)

table(merged_data$p_id)
table(merged_data$time)

rm(sleep_data_temp)
```

### Combining the time variables and creating a week variable

```{r}
merged_data$date_time <- with(merged_data, mdy(Date) + hms(Time))

merged_data$week_year <- week(merged_data$date_time)
merged_data$day <- day(merged_data$date_time)
```

### Creating physical activity variables 

I'm using a generalization of method from this paper  

Moran DS, Heled Y, Gonzalez RR. Metabolic rate monitoring and energy expenditure prediction using a novel actigraphy method. Med Sci Monit. 2004 Nov;10(11):MT117-20. Epub 2004 Oct 26. PMID: 15507861. [https://pubmed.ncbi.nlm.nih.gov/15507861/](https://pubmed.ncbi.nlm.nih.gov/15507861/)

- ZCM of 40 or greater is moderate to vigorous activity
- ZCM of 0 is sleep (**This is probably WRONG**)
- ZCM of sedentary to light activity is ZCM greater than 0 less than 40 

```{r}
merged_data <- merged_data %>% 
                mutate(activity_type = case_when(
                  ZCM >= 45 ~ "MVPA",
                  ZCM == 0 ~ "ZCM_Sleep",
                  ZCM > 0 & ZCM < 40 ~ "Sed_Light"
                ))

### Fix activity type for M_Sleep instead of ZCM sleep

merged_data <- merged_data %>% 
                mutate(activity_type_sleep = case_when(
                  M_Sleep == 1 ~ "sleep",
                  ZCM >= 45 ~ "mvpa",
                  ZCM < 45 ~ "sed_light"
                ))

table(merged_data$activity_type_sleep)

merged_data$M_Sleep <- as.numeric(merged_data$M_Sleep)
merged_data$TIB <- as.numeric(merged_data$TIB)
merged_data$SLP <- as.numeric(merged_data$SLP)
merged_data$ZCM <- as.numeric(merged_data$ZCM)
```

## Creating total time of week for summary activities

```{r}
merged_data <- merged_data %>%
	mutate(sed_light_mvpa_minutes = case_when(
		activity_type_sleep == "sed_light" ~ 1,
		activity_type_sleep == "sleep" ~ 0,
		activity_type_sleep == "mvpa" ~ 1
	))

merged_data <- merged_data %>%
	mutate(mvpa_minutes = case_when(
		activity_type_sleep == "sed_light" ~ 0,
		activity_type_sleep == "sleep" ~ 0,
		activity_type_sleep == "mvpa" ~ 1
	))

merged_data <- merged_data %>%
	mutate(sed_light_minutes = case_when(
		activity_type_sleep == "sed_light" ~ 1,
		activity_type_sleep == "sleep" ~ 0,
		activity_type_sleep == "mvpa" ~ 0
	))

merged_data <- merged_data %>%
	mutate(sleep = case_when(
		activity_type_sleep == "sed_light" ~ 0,
		activity_type_sleep == "sleep" ~ 1,
		activity_type_sleep == "mvpa" ~ 0
	))
```

### Summary Statistics - Sleep

```{r}
merged_data$total_sleep_mvpa_light <- merged_data$sleep + merged_data$mvpa_minutes + merged_data$sed_light_minutes

table(merged_data$total_sleep_mvpa_light)

summary(merged_data$total_sleep_mvpa_light)

merged_data$day <- as.factor(merged_data$day)

day <- merged_data %>%
  group_by(p_id, day, time) %>%
    summarise(sleep_minutes = sum(sleep),
              mvpa_minutes = sum(mvpa_minutes),
              sed_light_minutes = sum(sed_light_minutes),
              sed_light_mvpa_minutes = sum(sed_light_mvpa_minutes),
              total_sleep_mvpa_light = sum(total_sleep_mvpa_light),
              tib = sum(TIB),
              slp = sum(SLP),
              zcm = mean(ZCM),
              total_minutes = n()
              )

day$zcm <- as.integer(day$zcm)
summary(day$total_minutes)
```

## Plot of ZCM sleep versus SLEEP1 variable

```{r}
summary(day$sleep_minutes)
sleep_plot <- ggplot(day, aes(x = sleep_minutes)) +
              geom_histogram() +
              theme_classic()
plot(sleep_plot)

```

### Summary stats for all variables 

```{r}
day %>%
  group_by(time) %>%
    get_summary_stats(sleep_minutes)

day %>%
  group_by(time) %>%
    get_summary_stats(mvpa_minutes)

day %>%
  group_by(time) %>%
    get_summary_stats(sed_light_minutes)
```

### Checking daily MVPA and sleep overtime

### Plot for Sleep 
```{r}
sleep_plot <- ggplot(day, aes(x = day, y = sleep_minutes, group = time, colour = factor(time))) +
              geom_boxplot() +
              theme_classic()
plot(sleep_plot)
```

### Plot for MVPA
```{r}
mvpa_plot <- ggplot(day, aes(x = day, y = mvpa_minutes, group = time, colour = factor(time))) +
              geom_boxplot() +
              theme_classic()
plot(mvpa_plot)
```

### Plot for Sed, Light
```{r}
all_activity_plot <- ggplot(day, aes(x = day, y = sed_light_mvpa_minutes, group = time, colour = factor(time))) +
              geom_boxplot() +
              theme_classic()
plot(all_activity_plot)
```

# Aggregating to the day level 

```{r}
day_summary <- day %>%
  group_by(p_id, time) %>%
    summarise(sleep_minutes = median(sleep_minutes, na.rm = TRUE),
              mvpa_minutes = mean(mvpa_minutes, na.rm = TRUE),
              sed_light_minutes = mean(sed_light_minutes, na.rm = TRUE),
              sed_light_mvpa_minutes = mean(sed_light_mvpa_minutes, na.rm = TRUE),
              total_sleep_mvpa_light = mean(total_sleep_mvpa_light, na.rm = TRUE),
              tib = sum(tib, na.rm = TRUE),
              slp = sum(slp, na.rm = TRUE),
              )
```

# Writing CSV files

```{r}
write_csv(day, "participant_days.csv")
write_csv(day_summary, "day_summary.csv")
```

# Pivoting and writing to SAV (SPSS) format

```{}
day_only <- as.data.frame(day_only)

day_only_wide <- reshape(day_only, direction = "wide", idvar = "p_id", timevar = "time", v.names = c("sleep_minutes", "sleep_minutes_zcm", "mvpa_minutes", "sed_light_minutes", "sed_light_mvpa_minutes", "total_sleep_mvpa_light", "total_sleep_mvpa_light_zcm"), sep = "_")

day_only_wide$ParticipantNumber <- seq(1:97)

## SAV
write_sav(day_only_wide, "days_wide.sav")
```

